Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-27 intro.tex
lots of disorganized edits, mostly focused on justifying use of modalities, some comments for fixing in sec 4
intro tweaks

\paragraph{Contributions}
We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
implemented as an embedded separation logic within Iris~\cite{jung2018iris}.
The result is a separation logic that lifts a number of major semantic restrictions present in the few
prior logics tackling virtual address translation.
The logic we develop covers core reasoning principles for reasoning about memory configurations and code
reliant upon or manipulating those memory configurations in the presence of in-memory page tables, the primary
memory protection mechanism across Intel/AMD's x86-64 processors; ARM's application class processors including
AArch64 CPUs; as well as POWER, RISC-V, and other architectures.
We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
including:
switching address spaces, converting physical addresses to virtual addresses for
software page table walks, and complete page table walk which backbones other essential functionalities of
virtual-memory which we discuss for mapping and unmapping pages. These examples demonstrate the suitability and flexibility of
a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical
capabilities of prior logics, or revisits an example from prior work with more details (treating parts of the kernel
invariant never addressed by prior work) and fewer assumptions (e.g., verifying the reverse physical-to-virtual 
address conversion rather than assuming it has already occurred).

>>>>

\paragraph{Contributions}
We develop these ideas in the form of a logic for working with virtual-address-space-relative assertions,
implemented as an embedded separation logic within \replace{Iris}{\textsc{Iris}}~\cite{jung2018iris}.
The result is a separation logic that lifts a number of major semantic restrictions present in the few
prior logics tackling virtual address translation.
The logic we develop covers core reasoning principles for reasoning about memory configurations and code
reliant upon or manipulating those memory configurations in the presence of in-memory page tables, the primary
memory protection mechanism across Intel/AMD's x86-64 processors; ARM's application class processors including
AArch64 CPUs; as well as POWER, RISC-V, and other architectures.
We prove the soundness of our \textsf{vProp} logic with respect to a RISC-like fragment of \textsf{AMD64} instructions.
We verify simplified versions of several critical virtual-memory-related pieces of OS functionality, 
including:
switching address spaces, converting physical addresses to virtual addresses for
software page table walks, and \add{a} complete \add{software} page table walk \replace{which backbones}{ --- the backbone of} other essential functionalities of
virtual-memory which we discuss for mapping and unmapping pages. 
These examples demonstrate the suitability and flexibility of
a modal treatment of address spaces: each has a concise specification, each example either goes beyond the technical
capabilities of prior logics, or revisits an example from prior work
\replace{with more details (treating parts of the kernel
invariant never addressed by prior work) and fewer assumptions (e.g., verifying the reverse physical-to-virtual 
address conversion rather than assuming it has already occurred).}{
  in greater detail (e.g., verifying critical and challenging supporting code that was
  axiomatized or treated as out-of-scope in prior work).
}
